Privacy-Respecting Type Error Telemetry at Scale

### 2023-10-02 programming23/paper.tex
error density FS, +-200

% TODO BEFORE FINAL
% - uncomment acks
% - update all refs to artifacts to no longer be in future tense

@@

\author[a,b]{Ben Greenman}
\authorinfo{(\email{benjamin.l.greenman@gmail.com}) was a CIFellows 2020
postdoc at Brown Univesity and is now an assistant professor
at the recently-named Kahlert School of Computing at the University of Utah.}
\affiliation[a]{Brown University, Providence, RI, USA}
\affiliation[b]{University of Utah, Salt Lake City, UT, USA}
% \orcid{0000-0001-7078-9287}

@@

\begin{table}[t]\centering
  \caption{Size of analyzed code: number of files, number of lines, and lines in edit range}
  \label{t:codebase-size}
  % ;; (list* (max* vv) (median < vv) mm (stddev/mean mm vv) (percentile* vv)))
  % ;; percentile* = 0.95 -- 0.99
  % #hash((editrange . (1156036 926 3007043/817 30975.40553032358 (0.95 8220) (0.96 9858) (0.97 13656) (0.98 18956) (0.99 34725)))
  %       (event-count . (6079 138 38689/135 582.5752201972966 (0.95 960) (0.96 1174) (0.97 1304) (0.98 1836) (0.99 3302)))
  %       (files . (54884 7678 55257996/4639 11853.650098459446 (0.95 40029) (0.96 42771) (0.97 45417) (0.98 48688) (0.99 51761)))
  %       (lines . (1089963 3115 38928947/5992 22364.75846303099 (0.95 18561) (0.96 25892) (0.97 27014) (0.98 29725) (0.99 50547)))
  %       (timespan . (1387897376 845648 591343604248/185705 15724745.722577972 (0.95 10696064) (0.96 12715468) (0.97 15596940) (0.98 21153924) (0.99 35450460))))
  %% NOTE: 3 records have +1M lines, all from one nocheck session, all with 3 files
  %% - session "30,762,216,447,324"
  %% NOTE: 133 records have +52K files, all from one nocheck session
  %% - session "330,172,870,224,081" 
  %% (2nd biggest has 41K files "31,650,473,913,529")
%% Files has HUGE numbers thanks to data assets. Forget it.
% Files  &              &
% 11,911 & \stddev{12}  &
%  7,678 &              &
% 51,761 &              &
  \begin{tabular}[t]{lrrrrl}
    ~          & Mean  & Stddev & Median & \ptile{99} & Distribution \\
    Lines      & 6,497 & \stddev{22} & 3,115 & 50,547 & \includegraphics[width=0.14\columnwidth,valign=M]{img/lines-distribution.pdf} \\
    Edit Range & 3,680 & \stddev{31} & 926   & 34,725 & \includegraphics[width=0.14\columnwidth,valign=M]{img/editrange-distribution.pdf}
    %% TODO x-axis starts at 1, not zero
    %% TODO much bigger font or much smaller plot!!!
  \end{tabular}
\end{table}

@@

\begin{table}[t]\centering
  \caption{Session size in seconds and in number of records.}
  \label{t:session-size}
  %% NOTE max time = 1 million seconds = 15 days ...  +100 are over 2 days
  \begin{tabular}{lrrrrl}
    ~               & Mean & Stddev & Median & \ptile{99} & Distribution \\
    Time Span (sec) & 3,184 & \stddev{16} & 845 & 35,450 & \includegraphics[width=0.14\columnwidth,valign=M]{img/timespan-distribution.pdf} \\
    Record Count    & 286 & 583         & 138 & 3,302  & \includegraphics[width=0.14\columnwidth,valign=M]{img/event-count-distribution.pdf}
    %% TODO x-axis starts at 1
    %% TODO huge font, or smaller plot
  \end{tabular}
\end{table}

@@

    \begin{tabular}[t]{cc}
    %% TODO why text blurry??!
    Type errors & Background errors \\
    \labelbars{\includegraphics[width=0.20\columnwidth,valign=M]{img/error-by-mode-te.pdf}}
    & \labelbars{\includegraphics[width=0.20\columnwidth,valign=M]{img/error-by-mode-fs.pdf}}
  \end{tabular}
  \caption{Type and \FS{} errors grouped by mode.}
  \label{f:error-by-mode}
\end{figure}

@@

\begin{table}[t]\centering
  %% NOTE: popularity looks similar when [over all records] vs. [restricted to module switch records]
  \caption{Type error popularity for \mnonstrict{} and \mstrict{} modes.}
  %% {(all of the type errors in \mnocheck{} mode are \code{SyntaxError}s).}
  \label{t:type-error-count}
  \begin{tabular}{ll}
    \mnonstrict{} & \mstrict{} \\
    \begin{tabular}[t]{l@{~}r}
      \code{UnknownSymbol} & \pct{62.13} \\
      \code{SyntaxError} & \pct{15.42} \\
      \code{UnknownProperty} & \pct{8.28} \\
      \code{UnknownRequire} & \pct{3.13} \\
      \code{TypeMismatch} & \pct{2.44} \\
      \code{CountMismatch} & \pct{2.26} \\
      \code{OptionalValueAccess} & \pct{2.07} \\
      \code{GenericError} & \pct{2.06} \\
      \code{UnknownPropButGotLikeProp} & \pct{0.44} \\
      \code{CannotExtendTable} & \pct{0.42} \\
      \code{GenericExtraInformation} & \pct{0.32} \\
      \code{ModuleHasCyclicDependency} & \pct{0.23} \\
      \code{IllegalRequire} & \pct{0.20} \\
      \code{NotATable} & \pct{0.15} \\
      \code{CannotCallNonFunction} & \pct{0.15} \\
      \code{MissingProperties} & \pct{0.09} \\
      \code{FunctionExitsWithoutReturn} & \pct{0.07} \\
      \code{FunctionDoesNotTakeSelf} & \pct{0.07} \\
      \code{MissingUnionProperty} & \pct{0.02} \\
      \code{CannotInferBinaryOperation} & \pct{0.02} \\
      \code{OnlyTablesCanHaveMethods} & \pct{0.01} \\
      \code{DuplicateTypeDefinition} & {<\pct{0.01}} \\
      \code{TypesAreUnrelated} & {<\pct{0.01}} \\
    \end{tabular}
        &
    \begin{tabular}[t]{l@{~}r}
      \code{UnknownSymbol} & \pct{23.97} \\
      \code{TypeMismatch} & \pct{20.46} \\
      \code{UnknownProperty} & \pct{18.88} \\
      \code{SyntaxError} & \pct{9.31} \\
      \code{CannotInferBinaryOperation} & \pct{6.94} \\
      \code{MissingProperties} & \pct{4.04} \\
      \code{CountMismatch} & \pct{2.99} \\
      \code{OptionalValueAccess} & \pct{2.99} \\
      \code{FunctionExitsWithoutReturn} & \pct{2.99} \\
      \code{UnknownRequire} & \pct{2.37} \\
      \code{GenericError} & \pct{2.28} \\
      \code{NotATable} & \pct{1.32} \\
      \code{GenericExtraInformation} & \pct{0.35} \\
      \code{UnknownPropButGotLikeProp} & \pct{0.26} \\
      \code{IncorrectGenericParamCount} & \pct{0.18} \\
      \code{CannotCallNonFunction} & \pct{0.18} \\
      \code{IllegalRequire} & \pct{0.18} \\
      \code{ModuleHasCyclicDependency} & \pct{0.09} \\
      \code{CannotExtendTable} & \pct{0.09} \\
      \code{OccursCheckFailed} & \pct{0.09} \\
      \code{TypesAreUnrelated} & \pct{0.09} \\
    \end{tabular}
  \end{tabular}
\end{table}

@@

%\newpage
%
%\appendix
%
%\section{Data Details}
%
%Tips for interpreting the data:
%
%\begin{itemize}
%  \item
%    Edit ranges are non-negative because we
%    implemented the interval arithmetic for them using unsigned integers.
%    To filter out negative edit ranges, we removed all ranges
%    greater than 4 billion.
%
%  \item
%  \item
%  \item
%\end{itemize}
%
%\newpage

>>>>

% TODO BEFORE FINAL
% - explain edit range computations
% - uncomment acks
% - update all refs to artifacts to no longer be in future tense

@@

\author[a,b]{Ben Greenman}
\authorinfo{(\email{benjamin.l.greenman@gmail.com}) was a CIFellows 2020
postdoc at Brown University and is now an assistant professor in the
Kahlert School of Computing at the University of Utah.}
\affiliation[a]{Brown University, Providence, RI, USA}
\affiliation[b]{University of Utah, Salt Lake City, UT, USA}
% \orcid{0000-0001-7078-9287}

@@

\begin{table}[t]\centering
  \caption{Size of analyzed code: number of files, number of lines, and lines in edit range}
  \label{t:codebase-size}
  % ;; (list* (max* vv) (median < vv) mm (stddev/mean mm vv) (percentile* vv)))
  % ;; percentile* = 0.95 -- 0.99
  % #hash((editrange . (1156036 926 3007043/817 30975.40553032358 (0.95 8220) (0.96 9858) (0.97 13656) (0.98 18956) (0.99 34725)))
  %       (event-count . (6079 138 38689/135 582.5752201972966 (0.95 960) (0.96 1174) (0.97 1304) (0.98 1836) (0.99 3302)))
  %       (files . (54884 7678 55257996/4639 11853.650098459446 (0.95 40029) (0.96 42771) (0.97 45417) (0.98 48688) (0.99 51761)))
  %       (lines . (1089963 3115 38928947/5992 22364.75846303099 (0.95 18561) (0.96 25892) (0.97 27014) (0.98 29725) (0.99 50547)))
  %       (timespan . (1387897376 845648 591343604248/185705 15724745.722577972 (0.95 10696064) (0.96 12715468) (0.97 15596940) (0.98 21153924) (0.99 35450460))))
  %% NOTE: 3 records have +1M lines, all from one nocheck session, all with 3 files
  %% - session "30,762,216,447,324"
  %% NOTE: 133 records have +52K files, all from one nocheck session
  %% - session "330,172,870,224,081" 
  %% (2nd biggest has 41K files "31,650,473,913,529")
%% Files has HUGE numbers thanks to data assets. Forget it.
% Files  &              &
% 11,911 & \stddev{12}  &
%  7,678 &              &
% 51,761 &              &
  \begin{tabular}[t]{lrrrrl}
    ~          & Mean  & Stddev & Median & \ptile{99} & Distribution \\
    Lines      & 6,497 & \stddev{22} & 3,115 & 50,547 & \includegraphics[width=0.14\columnwidth,valign=M]{img/lines-distribution.pdf} \\
    Edit Range & 3,680 & \stddev{31} & 926   & 34,725 & \includegraphics[width=0.14\columnwidth,valign=M]{img/editrange-distribution.pdf}
    %% TODO x-axis starts at 1, not zero ... do we have zeros?! may need to explain
  \end{tabular}
\end{table}

@@

\begin{table}[t]\centering
  \caption{Session size in seconds and in number of records.}
  \label{t:session-size}
  %% NOTE max time = 1 million seconds = 15 days ...  +100 are over 2 days
  \begin{tabular}{lrrrrl}
    ~               & Mean & Stddev & Median & \ptile{99} & Distribution \\
    Time Span (sec) & 3,184 & \stddev{16} & 845 & 35,450 & \includegraphics[width=0.14\columnwidth,valign=M]{img/timespan-distribution.pdf} \\
    Record Count    & 286 & 583         & 138 & 3,302  & \includegraphics[width=0.14\columnwidth,valign=M]{img/event-count-distribution.pdf}
    %% TODO x-axis starts at 1
  \end{tabular}
\end{table}

@@

    \begin{tabular}[t]{cc}
    Type errors & Background errors \\
    \labelbars{\includegraphics[width=0.20\columnwidth,valign=M]{img/error-by-mode-te.pdf}}
    & \labelbars{\includegraphics[width=0.20\columnwidth,valign=M]{img/error-by-mode-fs.pdf}}
  \end{tabular}
  \caption{Type and \FS{} errors grouped by mode.}
  \label{f:error-by-mode}
\end{figure}

@@

\begin{table}[t]\centering
  %% NOTE: popularity looks similar when [over all records] vs. [restricted to module switch records]
  \caption{Type error popularity for \mnonstrict{} and \mstrict{} modes.}
  %% {(all of the type errors in \mnocheck{} mode are \code{SyntaxError}s).}
  \label{t:type-error-count}
  \begin{tabular}{l@{~~}l}
    \mnonstrict{} & \mstrict{} \\
    \begin{tabular}[t]{l@{~}r}
      \code{UnknownSymbol} & \pct{62.13} \\
      \code{SyntaxError} & \pct{15.42} \\
      \code{UnknownProperty} & \pct{8.28} \\
      \code{UnknownRequire} & \pct{3.13} \\
      \code{TypeMismatch} & \pct{2.44} \\
      \code{CountMismatch} & \pct{2.26} \\
      \code{OptionalValueAccess} & \pct{2.07} \\
      \code{GenericError} & \pct{2.06} \\
      \code{UnknownPropButGotLikeProp} & \pct{0.44} \\
      \code{CannotExtendTable} & \pct{0.42} \\
      \code{GenericExtraInformation} & \pct{0.32} \\
      \code{ModuleHasCyclicDependency} & \pct{0.23} \\
      \code{IllegalRequire} & \pct{0.20} \\
      \code{NotATable} & \pct{0.15} \\
      \code{CannotCallNonFunction} & \pct{0.15} \\
      \code{MissingProperties} & \pct{0.09} \\
      \code{FunctionExitsWithoutReturn} & \pct{0.07} \\
      \code{FunctionDoesNotTakeSelf} & \pct{0.07} \\
      \code{MissingUnionProperty} & \pct{0.02} \\
      \code{CannotInferBinaryOperation} & \pct{0.02} \\
      \code{OnlyTablesCanHaveMethods} & \pct{0.01} \\
      \code{DuplicateTypeDefinition} & {<\pct{0.01}} \\
      \code{TypesAreUnrelated} & {<\pct{0.01}} \\
    \end{tabular}
        &
    \begin{tabular}[t]{l@{~}r}
      \code{UnknownSymbol} & \pct{23.97} \\
      \code{TypeMismatch} & \pct{20.46} \\
      \code{UnknownProperty} & \pct{18.88} \\
      \code{SyntaxError} & \pct{9.31} \\
      \code{CannotInferBinaryOperation} & \pct{6.94} \\
      \code{MissingProperties} & \pct{4.04} \\
      \code{CountMismatch} & \pct{2.99} \\
      \code{OptionalValueAccess} & \pct{2.99} \\
      \code{FunctionExitsWithoutReturn} & \pct{2.99} \\
      \code{UnknownRequire} & \pct{2.37} \\
      \code{GenericError} & \pct{2.28} \\
      \code{NotATable} & \pct{1.32} \\
      \code{GenericExtraInformation} & \pct{0.35} \\
      \code{UnknownPropButGotLikeProp} & \pct{0.26} \\
      \code{IncorrectGenericParamCount} & \pct{0.18} \\
      \code{CannotCallNonFunction} & \pct{0.18} \\
      \code{IllegalRequire} & \pct{0.18} \\
      \code{ModuleHasCyclicDependency} & \pct{0.09} \\
      \code{CannotExtendTable} & \pct{0.09} \\
      \code{OccursCheckFailed} & \pct{0.09} \\
      \code{TypesAreUnrelated} & \pct{0.09} \\
    \end{tabular}
  \end{tabular}
\end{table}

@@

% \appendix
% 
% \section{Data Details}
% 
% Tips for interpreting the data:
% 
% \begin{itemize}
%   \item
%     A \code{"-"} entry denotes a zero.
%     This data was not in the telemetry record.
%     Kibana filled in the dash.
%   \item
%     Timestamps are California local time.
%     %% (Daylight savings?)
%   \item
%   \item
%     Edit ranges are non-negative because we
%     implemented the interval arithmetic for them using unsigned integers.
%     To filter out negative edit ranges, we removed all ranges
%     greater than 4 billion.
% 
%   \item
%   \item
% \end{itemize}
% 
% \section{More Plots}
% 
% \begin{figure}[t]\centering
%   %% +-200 background errors, same as ever! balanced
%   \mnocheck{}
%   \includegraphics[width=\columnwidth]{img/error-count-nocheck-row--fs-density-diff.pdf}
% 
%   \mnonstrict{}
%   \includegraphics[width=\columnwidth]{img/error-count-nonstrict-row--fs-density-diff.pdf}
% 
%   \mstrict{}
%   \includegraphics[width=\columnwidth]{img/error-count-strict-row--fs-density-diff.pdf}
%   \caption{Changes in \FS{} error density (errors / lines) over time (seconds).}
%   \label{f:error-density}
% \end{figure}
